---
title: "EDS 221 Day 2 Task"
format: html
editor: visual
execute:
  warning: false
---

## Day 2 Tasks / Activities

```{r}
library(tidyverse)
library(here)
library(janitor)
```

## Part 1. Checking data types

### Create some data, check the classes, index!

Vectors, lists & data frames

```{r}
vec_1 <- c(2, 5, 9, 10, 8, 12, 1, 0)
class(vec_1)
typeof(vec_1)

vec_1_e3 <- vec_1[3]
vec_1_e1 <- vec_1[1]
vec_1_e5to7 <- vec_1[5:7]

```

```{r}
vec_1_char <- as.character(vec_1)
class(vec_1_char)
```

```{r}
vec_2 <- c(town = "Santa Barbara", location = "Rincon", swell = "south")

class(vec_2)
length(vec_2)

vec_2_e2 <- vec_2[2]
```

```{r}
df_1 <- data.frame(region = c("A", "B", "A", "D"), species = c("otter", "great white", "sea lion", "gray whale"), count = c(12, 2, 36, 6))

class(df_1)

class(df_1$species)

max_count <- max(df_1$count)
```

## Part 2. Wild data

#### **Set-up**

-   Visit the EDI site to learn about Mack Creek salamander & cutthroat trout data you\'ll be using here: [**data package**](https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-and.4027.14)

-   Download the *first CSV* listed (`AS00601.csv`), and take a look at it (outside of R is fine as a first step, e.g. you can open the CSV in Excel)

-   Explore the metadata (see View Full Metadata in the [**Resources section**](https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-and.4027.14) of the data website)

-   What does each column contain? What are the units of each? What is the study overall about?
